---
title: 'Reproducible Research: Peer Assessment 1'
author: "Felipe Villota"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---



# Creating an setting the directory 

```{r}
dir()
if(!file.exists("R5P1")){dir.create("R5P1")}
dir("R5P1")
setwd("C:/Users/USER/Desktop/R/R5P1")
```



# Load and explore CSV file 

```{r}
DATA<- read.csv("activity.csv")
str(DATA)
head(DATA)
dim(DATA)
names(DATA)
```



# 1. What is mean total number of steps taken per day?

## Total number of steps taken per day

```{r}
stepsperday <- aggregate(DATA$steps, list(DATA$date), FUN=sum)
colnames(stepsperday) <- c("Date", "Steps")
stepsperday
```

## Histogram of the total number of steps taken each day

```{r}
library(ggplot2)
library(dplyr)
stephist <- ggplot(stepsperday, aes(Steps))
stephist+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", 
        fill="yellow")
        
```


## Mean and median of the total number of steps taken per day

```{r}
mean(stepsperday$Steps, na.rm=TRUE)

median(stepsperday$Steps, na.rm=TRUE)
```


# 2. What is the average daily activity pattern?

## Time series of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
stepsperint <- aggregate(steps~interval,data=DATA,FUN=mean,na.action=na.omit)
stepsperint$time <- stepsperint$interval/100
G <- ggplot(stepsperint, aes(time, steps))
G+geom_line(col="red")+ggtitle("Average steps per time interval")

```


## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
library(dplyr)
M <- tbl_df(stepsperint)
M %>% select(time, steps) %>% filter(steps==max(M$steps))
```


# 3. Imputing missing values

## Calculate and report the total number of missing values in the datase

```{r}
ACT <- tbl_df(DATA)
ACT %>% filter(is.na(steps)) %>% summarize(missing_values = n())
```


## Replace missing values
```{r}
DATA$CompleteSteps <- ifelse(is.na(DATA$steps), 
                        round(stepsperint$steps[match(DATA$interval, 
                        stepsperint$interval)],0), DATA$steps)
head(DATA$CompleteSteps)

## New dataset 

DATAFull <- data.frame(steps=DATA$CompleteSteps, 
                           interval=DATA$interval, date=DATA$date)


head(DATAFull)

```


##  Histogram of the total number of steps taken each day with missing data filled in
```{r}
Full <- aggregate(DATAFull$steps, list(DATAFull$date), FUN=sum)
colnames(Full) <- c("Date", "Steps")

H2 <- ggplot(Full, aes(Steps))
H2+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", fill="lightyellow")
      
```



## What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
mean(Full$Steps)
median(Full$Steps)

# Both decreased slightly

```

# 4. Are there differences in activity patterns between weekdays and weekends? 

##Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.



```{r}
DATAFull$RealDate <- as.Date(DATAFull$date, format = "%Y-%m-%d")
DATAFull$weekday <- weekdays(DATAFull$RealDate)



for(i in 1:length(DATAFull$date)){
  if(weekdays(as.Date(DATAFull$date[i]))=="Sábado"|weekdays(as.Date(DATAFull$date[i]))=="Domingo"){
    DATAFull$day[i]="weekend"
  }
  else{
    DATAFull$day[i]="weekday"
  }
}

head(DATAFull, n=10)

```



## Two time series plot of the 5-minute interval (x) and the average number of steps taken averaged across weekday days or weekend days (y).

```{r}
weekdata<-DATAFull %>% group_by(day,interval)%>% summarise(stepmean=mean(steps))
ggplot(weekdata,aes (interval, stepmean)) + geom_line() +facet_wrap(day~.,nrow=2,ncol=1)+ggtitle("Mean Steps by Interval depending on Day")
  
```


